The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2010-01-07 script/en/chapters/numbertheory.tex
Script[de+en]: Started to insert the chapter with the plots about the quantity of primes in higher ranges, and the according references; some very small fixes.

\begin{remark}{:}\\
The security of the RSA algorithm depends as with all public key methods
on the difficulty to calculate the private key $d$ from the public key $(n,e)$.
\end{remark}

\noindent Concretely for the RSA method does this mean:
\vskip - 1em
\begin{enumerate}
  \item It is hard to calculate $J(n)$ for big compounds $n$ and
  \item It is hard to calculate the prime factors of big compounds $n$
        \index{Factorization!factorization problem}.
\end{enumerate}
\vskip +1em

% --------------------------------------------------------------------------
\hypertarget{RSAproof}{}
\subsection{Proof of requirement 1 (invertibility)}
\label{RSAproof}

>>>>

\vskip +10pt
\begin{remark}{:}\\
The security of the RSA algorithm depends as with all public key methods
on the difficulty to calculate the private key $d$ from the public key $(n,e)$.
\end{remark}

\noindent Concretely for the RSA method does this mean:
\vskip - 1em
\begin{enumerate}
  \item It is hard to calculate $J(n)$ for big compounds $n$ and
  \item It is hard to calculate the prime factors of big compounds $n$
        (factorization problem).%
\footnote{%
    There is no reason for the concern sometimes mentioned that there are
    not enough primes: Raising the dimension (exponent) of the modul always
    offers enough primes to consider -- this is visualized in
    chapter~\ref{primes:_Appendix_Plotting-Primes-Quantity}
    ``\nameref{primes:_Appendix_Plotting-Primes-Quantity}''.
}
\index{Factorization!factorization problem}.
\end{enumerate}
\vskip +1em

% --------------------------------------------------------------------------
\vskip +10pt
\hypertarget{RSAproof}{}
\subsection{Proof of requirement 1 (invertibility)}
\label{RSAproof}

### 2010-01-07 script/en/chapters/paper_and_pencil.tex
Script[de+en]: Started to insert the chapter with the plots about the quantity of primes in higher ranges, and the according references; some very small fixes.

\item {\bf Rail Fence}\footnote{%
   This method can directly be found in CrypTool\index{CrypTool} at the menu item
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Scytale / Rail Fence}.
   You can simulate this method also under the menu {\bf Crypt/Decrypt \textbackslash{} 
   Symmetric (classic) \textbackslash{} Permutation}: For a Rail Fence with
   2 lines use as key ``B,A'' and accept the default settings (only one
   permutation, where your input is done line-by-line and the output is
   taken column-by-column). 
   Using the key ``A,B'' would start the zigzag pattern below in the
   way, that the first letter is written into the first line instead of the
   second line.}
   \cite{pp:Singh2001}\index{Rail Fence cipher}:
   The characters of a message are alternately written in two (or more) lines,
   creating a zigzag pattern. The resulting ciphertext is read out 
   line by line.\\
   This is more a children's method.

@@

\item {\bf Scytale}\footnote{%
   This method can directly be found in CrypTool\index{CrypTool} at the menu item
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Scytale / Rail Fence}.
   As this method is a special case of a simple columnar transposition, you also can
   simulate it in CrypTool\index{CrypTool} under the 
   menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
   Permutation}: For the Scytale within the dialog box only the first 
   permutation is used. If the wood has e.g.\ 4 angles use as key ``1,2,3,4''.
   This is equivalent to write the text horizontally in blocks of 4 letters 
   in a matrix and to read it out vertically . 
   Because the key is in an in ascending order, the Scytale is denoted as
   an identical permutation. And because writing and read-out is done only
   once it is a simple (and no double) permutation.}
   \cite{pp:Singh2001}\index{Scytale}%
   : 
   This method was probably used since 600 B.C. -- a description
   of how it operated is not known from before Plutarch (50-120 B.C.).\\
   A long strip of paper is wrapped around a wooden cylinder and then the 
   message is written along the length of this strip. The ciphertext is 
   produced by unwinding the strip.

@@

\item {\bf General monoalphabetic substitution / Random letter 
   pairs\footnotemark}
   \footnotetext{%
   This cipher can be simulated in CrypTool\index{CrypTool} 
   under the menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) 
   \textbackslash{} Substitution / Atbash}.} 
   \cite{pp:Singh2001}:
   The substitution occurs by a given assignment of single letters.

\item {\bf Atbash\footnotemark}
   \footnotetext{%
   This cipher can be simulated in CrypTool\index{CrypTool} 
   under the menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) 
   \textbackslash{} Substitution / Atbash}.}  
   \cite{pp:Singh2001}\index{Atbash}: Replace the first letter of the 
   alphabet by the last letter of the alphabet, the second one by the last 
   but one, etc.

\item {\bf Shift cipher, for example Caesar cipher}\footnote{In CrypTool
   this method  can be found at three different places in the menu tree:\\
   - {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
      Caesar / ROT13}\\
   - {\bf Analysis \textbackslash{} Symmetric Encryption (classic)
     \textbackslash{} Ciphertext only \textbackslash{} Caesar} \\
   - {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     \textbackslash{} Caesar}. } 
   \cite{pp:Singh2001}\index{Caesar}%
   : Plaintext alphabet and ciphertext alphabet are shifted against each 
   other by a determined number of letters. Using the Caesar cipher means 
   shifting letters about three positions.

@@

\item {\bf Substitution with symbols} \cite{pp:Singh2001}, for instance the 
   so-called ``freemason cipher'': Each letter is replaced with a symbol.

\item {\bf Variants}: Fill characters, intentional mistakes \cite{pp:Singh2001}.

@@

\item {\bf Coding} \cite{pp:Singh2001}: In the course of time, codebooks were
   used again and again. A codebook assigns a codeword, a symbol or a number
   to every possible {\bf word} of a message. Only if both parties hold
   identical codebooks and if the assignment of codewords to plaintext words
   is not revealed, a successful and secret communication can take place.

\item {\bf Nomenclature} \cite{pp:Singh2001}\index{Nomenclature}: 
   A nomenclature is an encryption 
   system that is based upon a ciphertext alphabet. This alphabet is used to 
   encrypt the bigger part of the message. Particularly frequent or top-secret
   words are replaced by a limited number of codewords existing besides the 
   ciphertext alphabet.

@@

\begin{itemize}
\item {\bf Homophonic monoalphabetic substitution}\footnote{This cipher can be 
   simulated in CrypTool\index{CrypTool} under the menu
   {\bf Crypt/Decrypt \textbackslash{}Symmetric (classic)\textbackslash{} Homophone}.}
   \cite{pp:Singh2001}: Each language has a typical frequency distribution of
   letters. To conceal this distribution, each plaintext letter is assigned
   several ciphertext characters.
   The number of ciphertext characters assigned depends on the frequency of the
   letter to be encrypted.

\item {\bf Beale cipher} \cite{pp:Singh2001}\index{Beale cipher}: 
   The Beale cipher is a book cipher
   that numbers the words of a keytext. These numbers replace the cleartext 
   letters by the words' initial letters.

@@

\begin{itemize}
\item {\bf ``Great Chiffre''} \cite{pp:Singh2001}: This cipher was used by 
   Louis XIV. and was not solved until the end of the nineteenth century. 
   Cryptograms consisted of 587 different numbers, every number representing
   a syllable. The inventors of the ``Great Chiffre'' (Rossignol, father and
   son) constructed additional traps to increase security.
   For example, a number could assign a different meaning to or delete the 
   preceding one.

   \hypertarget{playfair}{}
\item {\bf Playfair}\footnote{%
   In CrypTool\index{CrypTool} you can call this method under the menu
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Playfair}.}
   \cite{pp:Singh2001}\index{Playfair}:
   A 5x5-matrix is filled with the plaintext characters. For example, the
   different letters of a keyword are inserted first, followed by the
   remaining letters. The plaintext is divided into pairs, these digraphs
   are encrypted using the following rules:

@@

\begin{itemize}
\item {\bf Vigen\`ere}\footnote{%
   In CrypTool\index{CrypTool} you can call this method under the menu
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Vigen\`ere}.}  
   \cite{pp:Singh2001}\index{Vigen\`ere}: 
   Each plaintext character is encrypted
   with a different ciphertext alphabet that is determined by the characters of
   a keyword (the so-called Vigen\`ere-Tableau serves auxiliary means).
   If the plaintext is longer than the key, the latter is repeated.

@@

\item {\bf ADFG(V)X}\footnote{%
   In CrypTool\index{CrypTool} you can call this method under the menu
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
   ADFGVX}.}
   \cite{pp:Singh2001}\index{ADFGVX}%
   : 
   ADFG(V)X-encryption was developed in Germany during WW1. The alphabet 
   is filled into a 5x5 or 6x6 matrix, and columns and rows are marked with
   the letters ADFGX and V, depending on the size of the grille. Each 
   plaintext character is substituted by the corresponding pair of letters. 
   Finally, a (row-) transposition cipher is performed on the resulting text.

@@

\begin{itemize}
\item {\bf ``Pinprick encryption''} \cite{pp:Singh2001}: For centuries, this 
   simple encryption method has been put into practice for different reasons (actually steganography).
   During the Victorian Age, for example, small holes underneath letters in
   newspaper articles marked the characters of a plaintext, as sending
   a newspaper was much more cheaper than the postage on a letter.

@@

\bibitem[Bauer1995]{pp:Bauer1995} \index{Bauer 1995}
    Friedrich L. Bauer, \\
    {\em Entzifferte Geheimnisse}, Springer, 1995.

\bibitem[Bauer2000]{pp:Bauer2000} \index{Bauer 2000}
    Friedrich L. Bauer, \\
    {\em Decrypted Secrets}, Springer 1997, 2nd edition 2000.

@@

\bibitem[Hill1929]{pp:Hill1929} \index{Hill 1929}
   Lester S. Hill,\\
   ``Cryptography in an Algebraic Alphabet,''
   \emph{The American Mathematical Monthly}, 36(6):306--312, 1929.

\bibitem[Hill1931]{pp:Hill1931} \index{Hill 1931}
   Lester S. Hill,\\
   ``Concerning Certain Linear Transformation Apparatus of Cryptography,''
   \emph{The American Mathematical Monthly}, 38(3):135--154, 1931.

\bibitem[Nichols1996]{pp:Nichols1996} \index{Nichols 1996} 
    Randall K. Nichols, \\
    {\em Classical Cryptography Course, Volume 1 and 2}, \\
    Aegean Park Press 1996;
    or in 12 lessons online at \\
    \url{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}

\bibitem[Nguyen2009]{pp:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       See Sage publications
       \url{http://www.sagemath.org/library-publications.html}

@@

\bibitem[Schmeh2004]{pp:Schmeh2004}  \index{Schmeh 2004}
        Klaus Schmeh, \\
        {\em Die Welt der geheimen Zeichen. Die faszinierende Geschichte
        der Verschl\"usselung},\\ 
        W3L Verlag Bochum, 1. Auflage 2004.

\bibitem[Schmeh2007]{pp:Schmeh2007}  \index{Schmeh 2007}
        Klaus Schmeh, \\
        {\em Codeknacker gegen Codemacher. Die faszinierende Geschichte der Verschl\"usselung},\\ 
        W3L Verlag Bochum, 2. Auflage 2007.\\
	This is most current among the books dealing in a comprehensive manner
	with the history of cryptology. It contains a small collection of solved
	and unsolved crypto riddles. One of the challenges deals with a double
	column transposition using two long keys, which are differently in addition.

@@

\bibitem[Singh2001]{pp:Singh2001} \index{Singh 2001}
	Simon Singh, \\
	{\em Geheime Botschaften. Die Kunst der Verschl\"usselung von der 
        Antike bis in die Zeiten des Internet}, \\
	dtv, 2001.

@@

\end{thebibliography}

>>>>

\item {\bf Rail Fence}\footnote{%
   This method can directly be found in CrypTool\index{CrypTool} at the menu item
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Scytale / Rail Fence}.
   You can simulate this method also under the menu {\bf Crypt/Decrypt \textbackslash{} 
   Symmetric (classic) \textbackslash{} Permutation}: For a Rail Fence with
   2 lines use as key ``B,A'' and accept the default settings (only one
   permutation, where your input is done line-by-line and the output is
   taken column-by-column). 
   Using the key ``A,B'' would start the zigzag pattern below in the
   way, that the first letter is written into the first line instead of the
   second line.}
   \cite{pp:Singh1999}\index{Rail Fence cipher}:
   The characters of a message are alternately written in two (or more) lines,
   creating a zigzag pattern. The resulting ciphertext is read out 
   line by line.\\
   This is more a children's method.

@@

\item {\bf Scytale}\footnote{%
   This method can directly be found in CrypTool\index{CrypTool} at the menu item
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Scytale / Rail Fence}.
   As this method is a special case of a simple columnar transposition, you also can
   simulate it in CrypTool\index{CrypTool} under the 
   menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
   Permutation}: For the Scytale within the dialog box only the first 
   permutation is used. If the wood has e.g.\ 4 angles use as key ``1,2,3,4''.
   This is equivalent to write the text horizontally in blocks of 4 letters 
   in a matrix and to read it out vertically . 
   Because the key is in an in ascending order, the Scytale is denoted as
   an identical permutation. And because writing and read-out is done only
   once it is a simple (and no double) permutation.}
   \cite{pp:Singh1999}\index{Scytale}%
   : 
   This method was probably used since 600 B.C. -- a description
   of how it operated is not known from before Plutarch (50-120 B.C.).\\
   A long strip of paper is wrapped around a wooden cylinder and then the 
   message is written along the length of this strip. The ciphertext is 
   produced by unwinding the strip.

@@

\item {\bf General monoalphabetic substitution / Random letter 
   pairs\footnotemark}
   \footnotetext{%
   This cipher can be simulated in CrypTool\index{CrypTool} 
   under the menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) 
   \textbackslash{} Substitution / Atbash}.} 
   \cite{pp:Singh1999}:
   The substitution occurs by a given assignment of single letters.

\item {\bf Atbash\footnotemark}
   \footnotetext{%
   This cipher can be simulated in CrypTool\index{CrypTool} 
   under the menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) 
   \textbackslash{} Substitution / Atbash}.}  
   \cite{pp:Singh1999}\index{Atbash}: Replace the first letter of the 
   alphabet by the last letter of the alphabet, the second one by the last 
   but one, etc.

\item {\bf Shift cipher, for example Caesar cipher}\footnote{In CrypTool
   this method  can be found at three different places in the menu tree:\\
   - {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
      Caesar / ROT13}\\
   - {\bf Analysis \textbackslash{} Symmetric Encryption (classic)
     \textbackslash{} Ciphertext only \textbackslash{} Caesar} \\
   - {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     \textbackslash{} Caesar}. } 
   \cite{pp:Singh1999}\index{Caesar}%
   : Plaintext alphabet and ciphertext alphabet are shifted against each 
   other by a determined number of letters. Using the Caesar cipher means 
   shifting letters about three positions.

@@

\item {\bf Substitution with symbols} \cite{pp:Singh1999}, for instance the 
   so-called ``freemason cipher'': Each letter is replaced with a symbol.

\item {\bf Variants}: Fill characters, intentional mistakes \cite{pp:Singh1999}.

@@

\item {\bf Coding} \cite{pp:Singh1999}: In the course of time, codebooks were
   used again and again. A codebook assigns a codeword, a symbol or a number
   to every possible {\bf word} of a message. Only if both parties hold
   identical codebooks and if the assignment of codewords to plaintext words
   is not revealed, a successful and secret communication can take place.

\item {\bf Nomenclature} \cite{pp:Singh1999}\index{Nomenclature}: 
   A nomenclature is an encryption 
   system that is based upon a ciphertext alphabet. This alphabet is used to 
   encrypt the bigger part of the message. Particularly frequent or top-secret
   words are replaced by a limited number of codewords existing besides the 
   ciphertext alphabet.

@@

\begin{itemize}
\item {\bf Homophonic monoalphabetic substitution}\footnote{This cipher can be 
   simulated in CrypTool\index{CrypTool} under the menu
   {\bf Crypt/Decrypt \textbackslash{}Symmetric (classic)\textbackslash{} Homophone}.}
   \cite{pp:Singh1999}: Each language has a typical frequency distribution of
   letters. To conceal this distribution, each plaintext letter is assigned
   several ciphertext characters.
   The number of ciphertext characters assigned depends on the frequency of the
   letter to be encrypted.

\item {\bf Beale cipher} \cite{pp:Singh1999}\index{Beale cipher}: 
   The Beale cipher is a book cipher
   that numbers the words of a keytext. These numbers replace the cleartext 
   letters by the words' initial letters.

@@

\begin{itemize}
\item {\bf ``Great Chiffre''} \cite{pp:Singh1999}: This cipher was used by 
   Louis XIV. and was not solved until the end of the nineteenth century. 
   Cryptograms consisted of 587 different numbers, every number representing
   a syllable. The inventors of the ``Great Chiffre'' (Rossignol, father and
   son) constructed additional traps to increase security.
   For example, a number could assign a different meaning to or delete the 
   preceding one.

   \hypertarget{playfair}{}
\item {\bf Playfair}\footnote{%
   In CrypTool\index{CrypTool} you can call this method under the menu
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Playfair}.}
   \cite{pp:Singh1999}\index{Playfair}:
   A 5x5-matrix is filled with the plaintext characters. For example, the
   different letters of a keyword are inserted first, followed by the
   remaining letters. The plaintext is divided into pairs, these digraphs
   are encrypted using the following rules:

@@

\begin{itemize}
\item {\bf Vigen\`ere}\footnote{%
   In CrypTool\index{CrypTool} you can call this method under the menu
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Vigen\`ere}.}  
   \cite{pp:Singh1999}\index{Vigen\`ere}: 
   Each plaintext character is encrypted
   with a different ciphertext alphabet that is determined by the characters of
   a keyword (the so-called Vigen\`ere-Tableau serves auxiliary means).
   If the plaintext is longer than the key, the latter is repeated.

@@

\item {\bf ADFG(V)X}\footnote{%
   In CrypTool\index{CrypTool} you can call this method under the menu
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
   ADFGVX}.}
   \cite{pp:Singh1999}\index{ADFGVX}%
   : 
   ADFG(V)X-encryption was developed in Germany during WW1. The alphabet 
   is filled into a 5x5 or 6x6 matrix, and columns and rows are marked with
   the letters ADFGX and V, depending on the size of the grille. Each 
   plaintext character is substituted by the corresponding pair of letters. 
   Finally, a (row-) transposition cipher is performed on the resulting text.

@@

\begin{itemize}
\item {\bf ``Pinprick encryption''} \cite{pp:Singh1999}: For centuries, this 
   simple encryption method has been put into practice for different reasons (actually steganography).
   During the Victorian Age, for example, small holes underneath letters in
   newspaper articles marked the characters of a plaintext, as sending
   a newspaper was much more cheaper than the postage on a letter.

@@

\bibitem[Bauer1995]{pp:Bauer1995} \index{Bauer 1995}
    Friedrich L. Bauer, \\
    {\em Entzifferte Geheimnisse},\\
    Springer, 1995.

\bibitem[Bauer2000]{pp:Bauer2000} \index{Bauer 2000}
    Friedrich L. Bauer, \\
    {\em Decrypted Secrets}, Springer 1997,\\
    2nd edition 2000.

@@

\bibitem[Hill1929]{pp:Hill1929} \index{Hill 1929}
   Lester S. Hill,\\
   \emph{Cryptography in an Algebraic Alphabet},\\
   The American Mathematical Monthly, 36(6): 306--312, 1929.

\bibitem[Hill1931]{pp:Hill1931} \index{Hill 1931}
   Lester S. Hill,\\
   \emph{Concerning Certain Linear Transformation Apparatus of Cryptography},\\
   The American Mathematical Monthly, 38(3): 135--154, 1931.

\bibitem[Nichols1996]{pp:Nichols1996} \index{Nichols 1996} 
    Randall K. Nichols, \\
    {\em Classical Cryptography Course, Volume 1 and 2}, \\
    Aegean Park Press 1996;\\
    or in 12 lessons online at \\
    \url{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}

\bibitem[Nguyen2009]{pp:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       See Sage publications \url{http://www.sagemath.org/library-publications.html},\\
       \url{http://www.sagemath.org/files/thesis/nguyen-thesis-2009.pdf},\\
       \url{http://sites.google.com/site/nguyenminh2/honours-thesis-2009.pdf}

@@

\bibitem[Schmeh2004]{pp:Schmeh2004}  \index{Schmeh 2004}
        Klaus Schmeh, \\
        {\em Die Welt der geheimen Zeichen. Die faszinierende Geschichte
        der Verschl\"usselung},\\ 
        W3L Verlag Bochum, 1st edition 2004.

\bibitem[Schmeh2007]{pp:Schmeh2007}  \index{Schmeh 2007}
        Klaus Schmeh, \\
        {\em Codeknacker gegen Codemacher. Die faszinierende Geschichte der Verschl\"usselung},\\ 
        W3L Verlag Bochum, 2nd edition 2007.\\
	This is most current among the books dealing in a comprehensive manner
	with the history of cryptology. It contains a small collection of solved
	and unsolved crypto riddles. One of the challenges deals with a double
	column transposition using two long keys, which are differently in addition.

@@

\bibitem[Singh1999]{pp:Singh1999} \index{Singh 1999}
	Simon Singh, \\
	{\em The Code Book: The Science of Secrecy from Ancient Egypt to Quantum Cryptography}, \\
	Anchor, 1999.

@@

\end{thebibliography}  % English

### 2010-01-07 script/en/chapters/primes.tex
Script[de+en]: Started to insert the chapter with the plots about the quantity of primes in higher ranges, and the according references; some very small fixes.

\newpage
\hypertarget{Kapitel_2}{}
\chapter{Prime Numbers}
\label{Label_Kapitel_2}
(Bernhard Esslinger, May 1999; Updates Nov. 2000, Dec. 2001, June 2003, May 2005, March 2006, June 2007, January 2010)

@@

% --------------------------------------------------------------------------
\subsection{\texorpdfstring{$f(n) = n^2 - 79 \cdot n + 1,601$}{f(n) = n\^{}2 - 79 ... n + 1,601}}
\label{L-Polynomfunktion02-1601}   \label{SrcPrimArith1}
   This function\footnote{%
   See \hyperlink{Primes_Appendix_A}{Appendix A of this chapter} for the source code
   to compute the table using Sage\index{Sage}.}  %Pari-GP.\index{Pari-GP}}
   delivers prime numbers for all values from $n=0$ to $n=79$.
Unfortunately $f(80) = 1,681 = 11 \cdot 151$ is not a prime number. To this
date, no function has been found that delivers more prime numbers in a row. On
the other hand, each prime occurs twice (first in the decreasing then in the
increasing sequence), which means that the algorithm delivers a total of 40
different prime values (these are the same ones as delivered by the function
in chapter~\ref{L-Polynomfunktion01-41})\footnote{%
   Another quadratic polynom, which delivers these primes, is: 
   $f(n) = n^2 - 9 \cdot n + 61$.\\
   Among the first 1000 sequence elements more than 50\% are prime
   (See \hyperlink{Primes_Appendix_A}{Appendix A of this chapter}).}.

@@

The \textbf{distribution} of primes displays several irregularities for which
no ``system'' has yet been found: On the one hand, many occur closely together,
like $2$ and $3,$ $ 11$ and $13, $ $ 809$ and $811$, on the other hand large
gaps containing no primes also occur. For example, no primes lie between $113$
and $127, $ $ 293$ and $307, $ $317$ and $331, $ $ 523$ and $541, $ $ 773$
and $787, $ $ 839$ and $853$ as well as between $887$ and $907$. \\
For details, please see: 
\vspace{-10pt}
\begin{itemize}
  \item[] \url{http://www.utm.edu/research/primes/notes/gaps.html}
\end{itemize}

@@

\paragraph{Further interesting topics regarding prime numbers}
\mbox{}\\
This chapter \ref{Label_Kapitel_2} didn't consider other number theory topics
such as divisibility rules, modulus calculation, modular inverses, modular
powers, modular roots, Chinese remainder theorem, Euler Phi function or
perfect numbers.
Some of these topics are considered in the \hyperlink{Chapter_ElementaryNT} 
{{\bf next chapter}} (chapter \ref{Chapter_ElementaryNT}).

@@

% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\newpage
\hypertarget{Primes_Appendix_A}{}
\section{Appendix: Examples using Sage}
% \section*{Appendix A: Examples using Sage}
% \addcontentsline{toc}{section}{Appendix A: Examples using Sage}
\label{Primes_Sage_samples}
\index{Sage!Code examples}
\index{Sage}

\noindent Below is Sage source code related to contents of the
chapter~\ref{Label_Kapitel_2} (``\nameref{Label_Kapitel_2}'').

>>>>

\newpage
% \hypertarget{Kapitel_2}{}
\chapter{Prime Numbers}
\label{Label_Kapitel_Primes}
(Bernhard Esslinger, May 1999; Updates Nov. 2000, Dec. 2001, June 2003, May 2005, March 2006, June 2007, January 2010)

@@

% --------------------------------------------------------------------------
\subsection{\texorpdfstring{$f(n) = n^2 - 79 \cdot n + 1,601$}{f(n) = n\^{}2 - 79 ... n + 1,601}}
\label{L-Polynomfunktion02-1601}   \label{SrcPrimArith1}
   This function\footnote{%
   See chapter~\ref{primes:_Appendix_Sage-Samples} ``\nameref{primes:_Appendix_Sage-Samples}''
   for the source code
   to compute the table using Sage\index{Sage}.}  %Pari-GP.\index{Pari-GP}}
   delivers prime numbers for all values from $n=0$ to $n=79$.
Unfortunately $f(80) = 1,681 = 11 \cdot 151$ is not a prime number. To this
date, no function has been found that delivers more prime numbers in a row. On
the other hand, each prime occurs twice (first in the decreasing then in the
increasing sequence), which means that the algorithm delivers a total of 40
different prime values (these are the same ones as delivered by the function
in chapter~\ref{L-Polynomfunktion01-41})\footnote{%
   Another quadratic polynom, which delivers these primes, is: 
   $f(n) = n^2 - 9 \cdot n + 61$.\\
   Among the first 1000 sequence elements more than 50\% are prime
   (See chapter~\ref{primes:_Appendix_Sage-Samples} ``\nameref{primes:_Appendix_Sage-Samples}''.).}.

@@

The \textbf{distribution} of primes%
\footnote{%
    Some visualizations (plots) of the quantity of primes in different
    number dimensions can be found in
    chapter~\ref{primes:_Appendix_Plotting-Primes-Quantity}
    ``\nameref{primes:_Appendix_Plotting-Primes-Quantity}''.
} displays several irregularities for which
no ``system'' has yet been found: On the one hand, many occur closely together,
like $2$ and $3,$ $ 11$ and $13, $ $ 809$ and $811$, on the other hand large
gaps containing no primes also occur. For example, no primes lie between $113$
and $127, $ $ 293$ and $307, $ $317$ and $331, $ $ 523$ and $541, $ $ 773$
and $787, $ $ 839$ and $853$ as well as between $887$ and $907$. \\
For details, please see: 
\vspace{-10pt}
\begin{itemize}
  \item[] \url{http://www.utm.edu/research/primes/notes/gaps.html}
\end{itemize}

@@

\paragraph{Further interesting topics regarding prime numbers}
\mbox{}\\
This chapter \ref{Label_Kapitel_Primes} didn't consider other number theory topics
such as divisibility rules, modulus calculation, modular inverses, modular
powers, modular roots, Chinese remainder theorem, Euler Phi function or
perfect numbers.
Some of these topics are considered in the \hyperlink{Chapter_ElementaryNT} 
{{\bf next chapter}} (chapter \ref{Chapter_ElementaryNT}).

@@

% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\newpage
\hypertarget{primes:_Appendix_Plotting-Primes-Quantity}{}
\section{Appendix: Plotting the number of primes in higher ranges}
\label{primes:_Appendix_Plotting-Primes-Quantity}
\index{Sage!Code examples}
\index{Sage}

\noindent xxxxxxxxx.

% ---------------------------------------------------------------------------
\subsection{xxxxxxxxxxxx}

xxxxxxxxxxxxxxxxxxxxxx

% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\newpage
\hypertarget{primes:_Appendix_Sage-Samples}{}
\section{Appendix: Examples using Sage}
% \section*{Appendix A: Examples using Sage}
% \addcontentsline{toc}{section}{Appendix A: Examples using Sage}
\label{primes:_Appendix_Sage-Samples}
\index{Sage!Code examples}
\index{Sage}

\noindent Below is Sage source code related to contents of the
chapter~\ref{Label_Kapitel_Primes} (``\nameref{Label_Kapitel_Primes}'').
